---
title: "Pacman"
subtitle: "Project"
editor: visual
execute: 
  warning: false
  message: false
---

### Loading necessary libraries

```{r}
#| label: load-pkgs
#| message: false

# Load the necessary libraries
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
library(scales)
library(plotly)
```

### Data Initialization

```{r}
#| label: load-dataset
#| message: false
data <- read.csv("data/Different_stores_dataset.csv")
glimpse(data)

```

### Data Preprocessing

```{r}
#| label: data-preprocessing
#| message: false

sum(is.na(data))

# Check unique values for categorical columns
unique(data$category)
unique(data$region)
unique(data$segment)

# Convert order_date and ship_date to date format
data$order_date <- as.Date(data$order_date, format = "%m/%d/%Y")
data$ship_date <- as.Date(data$ship_date, format = "%m/%d/%Y")

# Extract Year, Month, and Season from order_date
data$year <- year(data$order_date)
data$month <- month(data$order_date)
data$season <- case_when(
  data$month %in% c(12, 1, 2) ~ "Winter",
  data$month %in% c(3, 4, 5) ~ "Spring",
  data$month %in% c(6, 7, 8) ~ "Summer",
  data$month %in% c(9, 10, 11) ~ "Fall"
)

# Calculate total sales and profit margins
data <- data %>%
  mutate(
    total_sales = sales * quantity,
    profit_margin = profit / sales * 100
  )

# Handle any missing or negative values
data <- data %>%
  filter(!is.na(total_sales) & !is.na(profit_margin))

# Group data by season, category, and region
aggregated_data <- data %>%
  group_by(season, category, region, segment) %>%
  summarise(
    avg_discount = mean(discount, na.rm = TRUE),
    total_sales = sum(total_sales, na.rm = TRUE),
    avg_profit_margin = mean(profit_margin, na.rm = TRUE),
    .groups = 'drop'
  )

# Normalize total_sales
data$total_sales_normalized <- (data$total_sales - min(data$total_sales, na.rm = TRUE)) /
                               (max(data$total_sales, na.rm = TRUE) - min(data$total_sales, na.rm = TRUE))

# Ensure categorical variables are treated as factors
data$category <- as.factor(data$category)
data$segment <- as.factor(data$segment)
data$region <- as.factor(data$region)
data$season <- factor(data$season, levels = c("Winter", "Spring", "Summer", "Fall"))
```

### Question 1: Seasonal Trends and Sales Performance:

#### Plot 1: Monthly sales by Category

```{r}
# Prepare dataâ 
monthly_sales <- data %>%
  group_by(month, category) %>%
  summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = 'drop')

# Create the enhanced plot with a legend title
ggplot(monthly_sales, aes(x = month, y = total_sales, color = category)) +
  geom_line(size = 1) +                             
  geom_point(size = 2) +                            
  geom_smooth(method = "lm", se = FALSE,         
              linetype = "solid", color = "black", linewidth = 0.2) +
  labs(
    title = "Monthly Sales Trends by Category",
    x = "Month",
    y = "Total Sales",
    color = "Product Category") +
  theme_minimal(base_size = 14) +                   
  theme(
    plot.title = element_text(hjust = 0.5,           
                              size = 16,
                              family = "Arial"),
    plot.title.position = "plot",
    axis.text.x = element_text(hjust = 0.5,
                               size = 12,
                               family = "Arial"),          
    axis.title = element_text(hjust = 0.5,
                               size = 14,
                               family = "Arial"),
    axis.text = element_text(size = 12,
                             hjust = 0.5,
                             family = "Arial"),           
    panel.grid.major.x = element_blank(),          
    panel.grid.major.y = element_line(color = "gray", 
                                      linetype = "dashed"), 
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)     
  ) +
  scale_color_brewer(palette = "Set2") +           
  scale_x_continuous(breaks = 1:12, labels = 1:12) + 
  scale_y_continuous(labels = comma)              



```

#### Plot 2: Seasonal sales variation by region

```{r}
#| warning: false
#| message: false

# Data preparation 
seasonal_sales_malls <- data %>%
  group_by(season, region) %>%
  summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = 'drop')

# Static plot with all enhancements
p <- ggplot(seasonal_sales_malls, aes(x = season, y = total_sales, fill = region)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  geom_text(aes(label = comma(total_sales)), 
            position = position_dodge(width = 0.9), vjust = -0.3, size = 3) +
  facet_wrap(~ region, scales = "fixed") +
  scale_fill_brewer(palette = "OrRd") +
  geom_text(aes(label = ifelse(total_sales > 50000, comma(total_sales), "")), 
            position = position_dodge(width = 0.9), vjust = -0.3, color = "black", size = 3.5) +
  labs(
    title = "Seasonal Sales Variations by Region", 
    x = "Seasons", 
    y = "Total Sales",
    fill = "Regions"
  ) +
  
  # Polished theme
  theme_classic() +
  theme(
    plot.title = element_text(size = 14, family = "Arial", hjust = 0.5),
    plot.title.position = "plot",
    axis.text = element_text(size = 10, family = "Arial", hjust = 0.5),
    axis.title = element_text(size = 12, family = "Arial", hjust = 0.5),
    legend.title = element_text(size = 12, family = "Arial", hjust = 0.5),
    legend.position = "top"
  )
interactive_plot <- ggplotly(p)
interactive_plot
```

#### Plot 3: Seasonal Sales by Categories

```{r}
# Bar plot for seasonal sales across categories
seasonal_category_sales <- data %>%
  group_by(season, category) %>%
  summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = 'drop')

ggplot(seasonal_category_sales, aes(x = season, y = total_sales, fill = category)) +
  geom_col(position = "dodge") +
  labs(
    title = "Seasonal Sales Across Categories",
    x = "Seasons",
    y = "Total Sales",
    fill = "Category"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "BuPu") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, family = "Arial"),
    text = element_text(family = "Arial", size = 12, hjust = 0.5),
    axis.text = element_text(size = 10, hjust = 0.5, family = "Arial"),
    axis.title = element_text(size = 12, hjust = 0.5, family = "Arial")
  ) +
  scale_y_continuous(labels = comma)
```

#### Plot 4: Heatmap of Seasonal Sales by Category and Region

```{r}
# Faceted heatmap plot for seasonal sales across regions and categories
faceted_data <- data %>%
  group_by(season, category, region) %>%
  summarise(total_sales = sum(total_sales, na.rm = TRUE), .groups = 'drop')

# Create the ggplot2 plot as before
ggplot_plot <- ggplot(faceted_data, aes(x = season, y = category, fill = total_sales)) +
  geom_tile() +
  facet_wrap(~region) +
  labs(title = "Heatmap of Seasonal Sales by Category and Region", x = "Season", y = "Category") +
  scale_fill_gradient(
    low = "white",
    high = "blue",
    name = "Total Sales",
    breaks = scales::pretty_breaks(n = 5),
    labels = scales::label_comma()
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, family = "Arial", size = 16),
  axis.title.x = element_text(hjust = 0.5, family = "Arial", size = 14),
  axis.title.y = element_text(hjust = 0.5, family = "Arial", size = 14),
  axis.text.x = element_text(hjust = 0.5, family = "Arial", size = 12),
  axis.text.y = element_text(hjust = 0.5, family = "Arial", size = 12),
  legend.title = element_text(hjust = 0.5, family = "Arial", size = 14))

# Convert the ggplot2 plot to a plotly object
plotly_plot <- ggplotly(ggplot_plot)

# Customize the plot
plotly_plot <- plotly_plot %>%
  layout(title = list(text = "Heatmap of Seasonal Sales by Category and Region", x = 0.5))

# Display the interactive plot
plotly_plot
```

### Question 2: Discounts, Sales Volumes, and Profit Margins

#### Plot 1: Relation between Discounts and sales

```{r}

data$discount <- data$discount * 100

# Scatter plot for discounts vs. sales volumes
ggplot(data, aes(x = discount, y = quantity)) +
  geom_point(alpha = 0.5, color = "blue") +
  labs(title = "Correlation Between Discounts and Sales Volumes", x = "Discount (%)", y = "Sales") +
  theme_minimal() +
  theme(
  plot.title = element_text(hjust = 0.5, family = "Arial", size = 16),
  axis.title.x = element_text(hjust = 0.5, family = "Arial", size = 14),
  axis.title.y = element_text(hjust = 0.5, family = "Arial", size = 14),
  axis.text.x = element_text(hjust = 0.5, family = "Arial", size = 12),
  axis.text.y = element_text(hjust = 0.5, family = "Arial", size = 12)
)

```

#### Plot 2: Impact of Discounts on Profit Margin by Category

```{r}
ggplot(data, aes(x = discount, y = profit_margin)) +
  geom_point(alpha = 0.7, color = "salmon", size = 2.5) + 
  geom_smooth(method = "lm", se = FALSE, color = "steelblue", size = 1) + 
  facet_wrap(~ category, scales = "free") + 
  labs(
    title = "Impact of Discounts on Profit Margins by Category",
    subtitle = "Faceted by category for detailed insights",
    x = "Discount (%)",
    y = "Profit Margin (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", family = "Arial"), 
    plot.subtitle = element_text(hjust = 0.5, size = 12, family = "Arial"), 
    axis.title = element_text(size = 12, family = "Arial", face = "italic"), 
    strip.text = element_text(size = 12, family = "Arial", face = "bold"), 
    panel.grid.major = element_line(color = "grey90", linetype = "dotted"), 
    panel.grid.minor = element_blank(), 
    plot.margin = margin(10, 10, 10, 10) 
  )
```

#### Plot 3: Correlation matrix

```{r}
# Create a correlation matrix
cor_matrix <- data %>%
  select(discount, quantity, profit_margin) %>%
  cor(use = "complete.obs")

# Visualize the correlation matrix as a heatmap
library(reshape2)
cor_melt <- melt(cor_matrix)

ggplot(cor_melt, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  labs(title = "Correlation Matrix: Discounts, Sales Volumes, and Profit Margins", x = "", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

#### Plot 4: 3D Scatter plot to visualize profit margin, discount and the Quantity sold

```{r}
library(plotly)

plot_ly(
  data = data,
  x = ~discount,
  y = ~quantity,
  z = ~profit_margin,
  type = 'scatter3d',
  mode = 'markers',
  marker = list(
    size = 5,
    color = ~profit_margin,       # Color points based on profit_margin
    colorscale = 'Viridis',       # Use Viridis color scale for better visibility
    opacity = 0.8                 # Add transparency for less clutter
  ),
  hoverinfo = 'text',             # Display additional information on hover
  text = ~paste(
    "Discount: ", discount, 
    "<br>Quantity: ", quantity, 
    "<br>Profit Margin: ", profit_margin
  )
) %>%
  layout(
    title = "3D Scatter Plot: Discounts vs Sales vs Profit Margins",
    scene = list(
      xaxis = list(
        title = "Discount (%)",
        showgrid = TRUE,
        zeroline = TRUE,
        backgroundcolor = 'rgb(240,240,240)',  # Light background for axis
        gridcolor = 'rgb(200,200,200)'         # Subtle grid lines
      ),
      yaxis = list(
        title = "Quantity Sold",
        showgrid = TRUE,
        zeroline = TRUE,
        backgroundcolor = 'rgb(240,240,240)',
        gridcolor = 'rgb(200,200,200)'
      ),
      zaxis = list(
        title = "Profit Margin (%)",
        showgrid = TRUE,
        zeroline = TRUE,
        backgroundcolor = 'rgb(240,240,240)',
        gridcolor = 'rgb(200,200,200)'
      ),
      camera = list(
        eye = list(x = 1.5, y = 1.5, z = 1.5)  # Set initial camera position
      )
    ),
    annotations = list(  # Add an example annotation
      list(
        x = 20, y = 50, z = 0.2,
        text = "High Discount, Low Profit",
        showarrow = TRUE,
        arrowhead = 4,
        arrowcolor = 'red',
        font = list(size = 12)
      )
    ),
    paper_bgcolor = 'rgba(240,240,240,1)',  # Light paper background
    plot_bgcolor = 'rgba(240,240,240,1)',  # Light plot background
    showlegend = TRUE                      # Enable legend
  )

```
