---
title: "E-commerce Sales Trends and Consumer Insights: A Data-Driven Analysis"
subtitle: "Proposal"
author: 
  - name: "Pacman"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "In order to identify trends in consumer behavior, sales patterns, and profitability, this project examines e-commerce transaction data from different shopping centers and geographical areas. We seek to determine seasonal sales patterns, high-value client groups, and the effects of discounts by examining variables such as consumer demographics, product categories, and payment methods. The knowledge acquired will contribute to our comprehension of customer decision-making processes and the effects of various sales tactics on performance in various contexts."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(dplyr)
library(lubridate)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
data <- read.csv("data/Different_stores_dataset.csv")
glimpse(data)

```

## **Data Description**

#### **Source of the Data**

This dataset comes from Kaggle and includes transaction records from different shopping malls and regions, designed to give a broad overview of consumer habits, sales patterns, and seasonal trends in e-commerce.

#### **Collection Information**

The data covers all authorizations from October 1, 2006, to March 31, 2024. During this period, any transaction extensions, such as those for Working Capital, were approved under original agreements made before a temporary halt in EXIM Bank’s authority. No new transactions were approved during that lapse; only extensions of previous agreements were processed. This dataset was transferred to Catalog.Data.gov in September 2023 after Data.EXIM.gov ceased operations. With its multi-year scope, the dataset offers a consistent view of transaction details, customer demographics, and location-specific data, making it well-suited for financial and trend analyses over time.

#### **Description of Observations**

This dataset includes various types of information, organized into the following categories:

**Transaction Details:** Each transaction record includes a timestamp, the total amount, and any discounts applied, which are useful for analyzing financial patterns and identifying seasonal shifts in sales.

**Product Information:** Each item sold is categorized, enabling an analysis of product trends within specific categories and helping to segment the data by product type.

**Customer Demographics:** Information on customers’ age and gender supports demographic analysis, helping to reveal spending preferences and behaviors across different customer groups.

**Regional Indicators:** Transactions are associated with specific shopping malls or regions, making it possible to compare how different locations perform.

**Payment Methods:** The dataset captures the type of payment used—such as credit card, mobile payment, or cash—which provides insights into payment preferences and how they may affect profit margins.

## **Ethical Considerations**

While the dataset includes customer demographic details like age and gender, it does not contain any personal identifiers, helping to maintain privacy. Since the dataset is publicly available on Kaggle, it appears any necessary anonymization was applied beforehand. However, careful interpretation is needed when using demographic data to avoid inadvertently reinforcing stereotypes or drawing unsupported conclusions about certain groups.

## **Data Preprocessing**

```{r}
#| label: data-preprocessing
#| message: false

# Checking for missing values
sum(is.na(data))

# Viewing unique categories, malls, and payment methods for verification
unique(data$category)
unique(data$shopping_mall)
unique(data$payment_method)

# Converting `invoice_date` to Date-Time format
data$invoice_date <- as.POSIXct(data$invoice_date, format = "%m/%d/%Y %H:%M")

# Extracting Year, Month, Day, and Hour for further analysis
data$year <- year(data$invoice_date)
data$month <- month(data$invoice_date)
data$day <- day(data$invoice_date)
data$hour <- hour(data$invoice_date)

# Defining seasons based on months
data$season <- case_when(
  data$month %in% c(12, 1, 2) ~ "Winter",
  data$month %in% c(3, 4, 5) ~ "Spring",
  data$month %in% c(6, 7, 8) ~ "Summer",
  data$month %in% c(9, 10, 11) ~ "Fall"
)

# Calculating Discount and Profit Margin
data <- data %>%
  mutate(
    discount = selling_price_per_unit - cost_price_per_unit,
    profit_margin = (selling_price_per_unit - cost_price_per_unit) / selling_price_per_unit * 100
  )

# Aggregating for seasonal sales performance
seasonal_data <- data %>%
  group_by(season, category, shopping_mall, payment_method) %>%
  summarise(
    total_sales = sum(selling_price_per_unit * quantity),
    avg_discount = mean(discount),
    avg_profit_margin = mean(profit_margin),
    .groups = 'drop'
  )
  

# Checking for outliers in discounts
boxplot(data$discount, main = "Boxplot for Discounts")

# Checking for outliers in profit margins
boxplot(data$profit_margin, main = "Boxplot for Profit Margins")

# Normalizing sales volume if necessary (optional)
data$total_sales_normalized <- (data$selling_price_per_unit * data$quantity - min(data$selling_price_per_unit * data$quantity)) / 
                               (max(data$selling_price_per_unit * data$quantity) - min(data$selling_price_per_unit * data$quantity))


# Ensuring factor levels for categorical variables
data$season <- factor(data$season, levels = c("Winter", "Spring", "Summer", "Fall"))
data$category <- as.factor(data$category)
data$shopping_mall <- as.factor(data$shopping_mall)
data$payment_method <- as.factor(data$payment_method)

```

## Questions

-   How Do Seasonal Trends Affect Sales Performance Across Different Categories, Malls, and Payment Methods?

    We are going to investigate the impact of seasonal fluctuations on sales across product categories, locations, and payment methods. This question will help identify peak seasons and preferred payment options during high-demand periods.

-   What Are the Relationships Between Discounted Prices, Sales Volumes, and Profit Margins in Different Shopping Malls and Product Categories?

    Through this question, we will examine the influence of discounts on sales volumes and profit margins, analyzing correlations across product categories and shopping malls.

## Analysis Plan

-   Seasonal Trends and Sales Performance:\

    1.  Analyze monthly sales trends by category.\
    2.  Segment data by malls to explore seasonal sales variations.\
    3.  Examine the distribution of payment methods by season.\
    4.  Compare seasonal sales across categories, payment methods, and malls.\
        Planned Visualizations: Line graphs, stacked area charts, heatmaps, and faceted bar plots.

-   Discounts, Sales Volumes, and Profit Margins:\

    1.  Calculate and analyze discount percentages.\
    2.  Assess correlation of discounts with sales volumes.\
    3.  Determine profit margins post-discounts.\
    4.  Conduct correlation analysis among discounts, volumes, and margins.\
        Planned Visualizations: Scatter plots, bar charts, correlation matrices, and 3D scatter plots.

## Citations

KZMontage, "Sales from different stores," Kaggle. \[Online\]. Available: [Sales from different stores.](https://www.kaggle.com/datasets/kzmontage/sales-from-different-stores).
